---
title: "JSON, XML, HTML"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    theme: yeti
    orientation: rows
    vertical_layout: fill
    
---

```{r}
library(rjson)
library(dplyr)
library(jsonlite)
library(XML)
library(RCurl)
library(compare)

```
Data 607 Assignment_7
==============

Inputs {.sidebar data-width=250}
-----------------------------------------------------------------------

### Assignment 7
### Anthony Munoz

Row {.tabset .tabset-fade}
-----------

### json
```{r echo=T}

df.json <- fromJSON("https://raw.githubusercontent.com/Anth350z/DATA-607/master/Assignment_6/Book_list.json")

df.json
```



### HTML
```{r echo=T}
data <- getURLContent("https://raw.githubusercontent.com/Anth350z/DATA-607/master/Assignment_6/Book_list.html")
df.html <- readHTMLTable(data)

df.html
```
### XML
```{r echo=T}
data <- getURLContent("https://raw.githubusercontent.com/Anth350z/DATA-607/master/Assignment_6/Book_list.xml")

df.xml <- xmlToDataFrame(data)
df.xml 
```

### Data Analysis

On this Assignment we worked on reading data from 3 different format JSON, XML, and HTML and after we should to some comparing analysis between.


```{r echo=T}
#JSON
str(df.json)
#HTML
str(df.html)
#XML
str(df.xml)

#using the compare library to see if the values on the dataframe are equal

compare(df.html,df.json)
compare(df.html,df.xml)
compare(df.xml,df.json)


```

when we see the data its seem very similar but when we use the compare function, this returns false maybe due to some little change that can be observed such title column name different and the data frame structure.
