---
title: 'DATA-607-Homework #1- Loading Data into a Data Frame'
author: "Anthony Munoz"
date: "2/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "DATA-607-Homework #1- Loading Data into a Data Frame "
output: html_document
author: "Anthony Munoz" 
date: "2/1/2019"
---

Data source: https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data <br / >
    (a) Mushroom records drawn from The Audubon Society Field Guide to North <br />
        American Mushrooms (1981). G. H. Lincoff (Pres.), New York: Alfred <br />
        A. Knopf <br />
    (b) Donor: Jeff Schlimmer (Jeffrey.Schlimmer@a.gp.cs.cmu.edu) <br />
    (c) Date: 27 April 1987 <br />
    retrieved: 2/1/2019
    
<h4>Data Dictionary</h4>

 <h5>Attribute Information: </h5>
 <div class= "row">
     1. classes:                  edible=e, poisonous=p 
     2. cap-shape:                bell=b,conical=c,convex=x,flat=f, 
                                  knobbed=k,sunken=s 
     3. cap-surface:              fibrous=f,grooves=g,scaly=y,smooth=s
     4. cap-color:                brown=n,buff=b,cinnamon=c,gray=g,green=r,
                                  pink=p,purple=u,red=e,white=w,yellow=y 
     5. bruises?:                 bruises=t,no=f 
     6. odor:                     almond=a,anise=l,creosote=c,fishy=y,foul=f, 
                                  musty=m,none=n,pungent=p,spicy=s 
     7. gill-attachment:          attached=a,descending=d,free=f,notched=n 
     7. gill-spacing:             close=c,crowded=w,distant=d 
     9. gill-size:                broad=b,narrow=n 
     10. gill-color:               black=k,brown=n,buff=b,chocolate=h,gray=g, 
                                  green=r,orange=o,pink=p,purple=u,red=e, 
                                  white=w,yellow=y 
    11. stalk-shape:              enlarging=e,tapering=t 
    12. stalk-root:               bulbous=b,club=c,cup=u,equal=e, 
                                  rhizomorphs=z,rooted=r,missing=? 
    13. stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s 
    14. stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s 
    15. stalk-color-above-ring:   brown=n,buff=b,cinnamon=c,gray=g,orange=o, 
                                  pink=p,red=e,white=w,yellow=y 
    16. stalk-color-below-ring:   brown=n,buff=b,cinnamon=c,gray=g,orange=o, 
                                  pink=p,red=e,white=w,yellow=y 
    17. veil-type:                partial=p,universal=u 
    18. veil-color:               brown=n,orange=o,white=w,yellow=y 
    19. ring-number:              none=n,one=o,two=t 
    20. ring-type:                cobwebby=c,evanescent=e,flaring=f,large=l, 
                                  none=n,pendant=p,sheathing=s,zone=z 
    21. spore-print-color:        black=k,brown=n,buff=b,chocolate=h,green=r, 
                                  orange=o,purple=u,white=w,yellow=y 
    22. population:               abundant=a,clustered=c,numerous=n, 
                                  scattered=s,several=v,solitary=y 
    23. habitat:                  grasses=g,leaves=l,meadows=m,paths=p, 
                                  urban=u,waste=w,woods=d 

</div>
<h4>Retrieving the Data</h4>



```{r}
##getting and saving the data.
data = read.csv(file = "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data", header = FALSE, sep = ",")
## data preview
head(data)
```

<h4>Subset</h4>

```{r}
## a subset from the and seelcting 5 variables  
data.frame = subset(data, select = c(V1,V2,V4,V9,V11))
## chaging the colunms name
 colnames(data.frame) <- c("class","cap-shape","cap-color","gill-size","stalk-shape")
 ## data preview
 head(data.frame)

```


<h4>renamimg the abbrebiation </h4>

```{r}
## selecting the abbreviation and replacing for names in each column
levels(data.frame$class) <- c("edible","poisonous")
levels(data.frame$`cap-shape`) <- c("bell","conical","flat","knobbed","sunken","convex")
levels(data.frame$`cap-color`) <- c("buff","cinnamon","red","gray","brown","pink","green","purple","white","yellow")
levels(data.frame$`gill-size`) <- c("broad","narrow")
levels(data.frame$`stalk-shape`) <- c("enlarging","tapering")
## data preview
head(data.frame)
## data summary
summary(data.frame)
```
