---
title: "NYC Restaurants Inspection"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    theme: yeti
    orientation: rows
    vertical_layout: fill
    
---

```{r setup, include=FALSE}
library(flexdashboard)
library(readr)
library(leaflet)
library(zipcode)
library(dplyr)
library(tidyr)
library(plotly)
library(DT)

library(tidyr)
library(treemapify)
library(stringr)
library(treemap)
library(revgeo)
library(tm)
library(wordcloud2)
library(RSocrata)
```


```{r}
#loading zipcode dataset
data("zipcode")

#retriving the data using RSocrata API
df <- read.socrata("https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j")

names(df) <- sub("\\.", " ", names(df))
word <- df


#retriving the violation code with the description
description <- df %>% group_by(`VIOLATION DESCRIPTION`,`VIOLATION CODE`) %>% select(`VIOLATION CODE`,`VIOLATION DESCRIPTION`) %>% unique() %>% na.omit()

#selecting the column needed and removing all the NAs
df <- as.data.frame(df) %>% select(BORO,ZIPCODE,`CUISINE DESCRIPTION`,`VIOLATION CODE`,`CRITICAL FLAG`,SCORE,GRADE, `INSPECTION DATE`)  %>% drop_na() 
#selecting the zipcode of both dataset to further join them
zip.code <- zipcode %>% select(ZIPCODE = zip, longitude, latitude) %>% na.omit 
#zip.code$ZIPCODE <- as.numeric(zip.code$ZIPCODE)

#used a left join by zipcode to in order to retrieve the longituve and latitude from zipcode dataset
df.new <- df %>% left_join(zip.code, by = "ZIPCODE") %>% drop_na() %>% filter(ZIPCODE != "12345") 


```
Overview & Description
===========================================
Inputs {.sidebar data-width=400}
-----------------------------------------------------------------------
### Anthony Munoz

On this project, I worked on NYC Restaurant inspections dataset. Some of the analysis I worked on it are: calculating and comparing violations data by year, weekdays and borough. Answer question such a witch year have the highest or lowest violations. 


I have used RSocrata API package to retrieve the data from the NYC OpenData website.

I retrieved the data for this project from NYC OpenData
link: https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j



Row {data-width=400}
-----------------------------------------------------------------------

### ok


```{r}

word.new <- word %>% group_by(`CUISINE DESCRIPTION`) %>%
     summarise(total = n())



wordcloud2(word.new,size=0.5,backgroundColor = "white" )

#graph for the all violation count


```



Dashboard 
====================

Row
-----------------------------------------------------------------------

### All Violations count

```{r}
#all the violation count of the 5 borough
valueBox(length(df.new$`VIOLATION CODE`),
         icon = "fas fa-book")

```

### Bronx

```{r}
#all the violation count for The Bronx
valueBox(length(df.new$BORO[df.new$BORO == "BRONX"] ),
         icon = 'fas fa-building')

```

### Brooklyn

```{r}
#all the violation count for Brooklyn
valueBox(length(df.new$BORO[df.new$BORO == "BROOKLYN"] ),
         icon = 'fas fa-building')

```


### Manhattan

```{r}
#all the violation count for Manhattan
valueBox(length(df.new$BORO[df.new$BORO == "MANHATTAN"] ),
         icon = 'fas fa-building')

```




### Queen

```{r}
#all the violation count for Queens
valueBox(length(df.new$BORO[df.new$BORO == "QUEENS"] ),
         icon = 'fas fa-building')

```


### Staten Island

```{r}
#all the violation count for Staten Island
valueBox(length(df.new$BORO[df.new$BORO == "STATEN ISLAND"] ),
         icon = 'fas fa-building')

```

Row
-------------------------------

### Violation Count

```{r}
#graph for the all violation count
pl <- df.new %>% group_by(`VIOLATION CODE`) %>%
        summarise(total = n()) %>%
          plot_ly(x=~`VIOLATION CODE` , 
                  y = ~total , 
                  type = "bar",
                   xaxis = list(autotick = T, dtick = 1),
                
                 marker=list(color= ~`VIOLATION CODE` , size=20 , opacity=0.9) ) %>% layout(xaxis = list(title = "Violation Code"),yaxis = list(title = "Data Count"))

pl


```


> This graph its showing the violations count. and we can see that the most common violation is "10F" which is "Non-food contact surface improperly constructed. Unacceptable material used. Non-food contact surface or equipment improperly maintained and/or not properly sealed, raised, spaced or movable to allow accessibility for cleaning on all sides, above and underneath the unit."

###  Violation count by borough

```{r}
#graph for the violation count by borough
pl <- df.new %>% group_by(BORO,.drop = T) %>%
        filter(BORO != "Missing" ) %>%
        summarise(total = n()) %>%
          plot_ly(x=~ BORO , 
                  y = ~total , 
                  type = "bar",
                 marker=list(color= ~total , size=20 , opacity=0.9)) %>% layout(xaxis = list(title = "Borough"),
                                                yaxis = list(title = "Violation Count"))

pl
```

Data Analysis 
====================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Violations by Weekday and Borough

````{r}
#extract the Days of the week and agragate it to a column on the dataset
df.new <- df.new %>% mutate(DAY = format(`INSPECTION DATE`, "%A"))

#group and get the total of count of week and borough exclude some "Missing" values
total.weekday.borough <- df.new %>% group_by(BORO,DAY) %>% summarise(total = n()) %>% filter(BORO != "Missing" ) 

#create a ggplot grapgh
gplot <- ggplot(total.weekday.borough, aes(x=BORO, y=total)) + geom_bar(stat = "identity", aes(fill=DAY), position ="dodge") + labs(x= "Borough", y="Violations Count ", title = "Violation by weekday and Borough")
#convert this ggplot object to a plotly object and get a dynamics graph
pl <- ggplotly(gplot)
pl
```


> On this graph we are comparing the weekday violations and Borough. we can see that Manhattan as the hightest number of violations.


### Violations by Year

````{r}
#this is the graph for the year counts
pl <- df.new %>% mutate( year = format(`INSPECTION DATE`, "%Y")) %>%
    group_by(year) %>%
        summarise(total = n()) %>%
          plot_ly(x=~ year , 
                  y = ~total , 
                  type = "bar",
                 marker=list(color= ~year , size=20 , opacity=0.5) ) %>% layout(xaxis = list(title = "Years"),yaxis = list(title = "Data Count"))
  

pl


```


> This is how many violations inspections happened by Year counts.we can see by looking at the graph that the numbers are increasing by year.



### Violations by Category

````{r}
#This is the graph for the category of critical or not critical
pl <- df.new %>% 
    group_by(`CRITICAL FLAG`) %>%
        summarise(total = n()) %>% na.omit() %>%
          plot_ly(x=~ `CRITICAL FLAG` , 
                  y = ~total , 
                  type = "bar",
                 marker=list(color= ~total , size=20 , opacity=0.8) ) %>% layout(xaxis = list(title = "Category"),yaxis = list(title = "Critical/Not Critical/Not Aplicable"))
  

pl


```

### Conclusion

The first thing that I saw showing on the graph was that yearly the restaurants are committing more violations being the most popular code violation "10F" which is "Non-food contact surface improperly constructed. Unacceptable material used. Non-food contact surface or equipment improperly maintained and/or not properly sealed, raised, spaced or movable to allow accessibility for cleaning on all sides, above and underneath the unit.". also I can see that the borough with most violation committed in Manhattan.


Map Marker
================

Row
----------------------------

### Map

```{r}

  #reduced the dataset for the map in order to get better performance
  leaflet(data = df.new[1:20000,]) %>% addTiles() %>%
    addMarkers(~longitude, ~latitude, label= ~as.character(`CRITICAL FLAG`), popup    = ~as.character(paste("Grade: ",GRADE,"Violation Code:", `VIOLATION CODE`,"Date:",`INSPECTION DATE`)),clusterOptions = markerClusterOptions() , options = leafletOptions(preferCanvas = TRUE) ) 

```

>For the Map markers I reduced the rows to 20,000 in order to get a better performance.

 Data Table
=========================

Row {.tabset .tabset-fade}
-------------------

### Data Table 

```{r}
#showing some of the data from the whole dataset. had to reduced the size of the data to obtain a better performace loading the data into datatable.
datatable(df.new[1:20000,])

```

### Dictionary for Violation Code

```{r}

datatable(description)
```





