---
title: "NYC Restaurants Inspection"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    theme: yeti
    orientation: rows
    vertical_layout: fill
    
---

```{r setup, include=FALSE}
library(flexdashboard)
library(readr)
library(leaflet)
library(zipcode)
library(dplyr)
library(tidyr)
library(plotly)
library(DT)

library(tidyr)
library(treemapify)
library(stringr)
library(treemap)
library(revgeo)
library(tm)
library(wordcloud2)
library(RSocrata)
library(wordcloud)
```


```{r}
#loading zipcode dataset
data("zipcode")

prueba <- read.socrata("https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j")


#retriving the data using RSocrata API
df <- read.socrata("https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j") %>% drop_na()

names(df) <- sub("\\.", " ", names(df))
word <- df


#retriving the violation code with the description
description <- df %>% group_by(`VIOLATION DESCRIPTION`,`VIOLATION CODE`) %>% select(`VIOLATION CODE`,`VIOLATION DESCRIPTION`) %>% unique() %>% na.omit()

#selecting the column needed and removing all the NAs
df <- as.data.frame(df) %>% select(BORO,ZIPCODE,`CUISINE DESCRIPTION`,`VIOLATION CODE`,`CRITICAL FLAG`,SCORE,GRADE, `INSPECTION DATE`)  %>% drop_na() 
#selecting the zipcode of both dataset to further join them
zip.code <- zipcode %>% select(ZIPCODE = zip, longitude, latitude) %>% na.omit 
#zip.code$ZIPCODE <- as.numeric(zip.code$ZIPCODE)

#used a left join by zipcode to in order to retrieve the longituve and latitude from zipcode dataset
df.new <- df %>% left_join(zip.code, by = "ZIPCODE") %>% drop_na() %>% filter(ZIPCODE != "12345") 


```

Overview & Description
===========================================


Inputs {.sidebar data-width=350}
-----------------------------------------------------------------------

### Anthony Munoz

This project explores the NYC Restaurant inspections dataset. The analysis includes exploring, calculating and comparing violations by year, weekdays and borough. This project answers questions such a which year has the highest or lowest number of violations.

I have used RSocrata API package to retrieve the data from the NYC OpenData website.

I retrieved the data for this project from NYC OpenData link:
 https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j



Column {data-width=650}
-----------------------------------------------------------------------

### Wordcloud


```{r}

word.cloud <- Corpus(VectorSource(word$`CUISINE DESCRIPTION`))
word.cloud <- tm_map(word.cloud, removePunctuation)

wordcloud(word.cloud, max.words = 2000, random.order = FALSE, col=colors(length(word.cloud)))



```



Dashboard 
====================

Row
-----------------------------------------------------------------------

### Number of Violations

```{r}
#all the violation count of the 5 borough
valueBox(length(df.new$`VIOLATION CODE`),
         icon = "fas fa-book")

```

### Bronx

```{r}
#all the violation count for The Bronx
valueBox(length(df.new$BORO[df.new$BORO == "BRONX"] ),
         icon = 'fas fa-building')

```

### Brooklyn

```{r}
#all the violation count for Brooklyn
valueBox(length(df.new$BORO[df.new$BORO == "BROOKLYN"] ),
         icon = 'fas fa-building')

```


### Manhattan

```{r}
#all the violation count for Manhattan
valueBox(length(df.new$BORO[df.new$BORO == "MANHATTAN"] ),
         icon = 'fas fa-building')

```




### Queen

```{r}
#all the violation count for Queens
valueBox(length(df.new$BORO[df.new$BORO == "QUEENS"] ),
         icon = 'fas fa-building')

```


### Staten Island

```{r}
#all the violation count for Staten Island
valueBox(length(df.new$BORO[df.new$BORO == "STATEN ISLAND"] ),
         icon = 'fas fa-building')

```

Row
-------------------------------

### Number of Violations by violation code

```{r}
#graph for the all violation count
pl <- df.new %>% group_by(`VIOLATION CODE`) %>%
        summarise(total = n()) %>%
          plot_ly(x=~`VIOLATION CODE` , 
                  y = ~total , 
                  type = "bar",
                   xaxis = list(autotick = T, dtick = 1),
                
                 marker=list(color= ~`VIOLATION CODE` , size=20 , opacity=0.9) ) %>% layout(xaxis = list(title = "Violation Code"),yaxis = list(title = "Number of Violations"))

pl


```


> This graph its showing the violations count. and we can see that the most common violation is "10F" which is "Non-food contact surface improperly constructed. Unacceptable material used. Non-food contact surface or equipment improperly maintained and/or not properly sealed, raised, spaced or movable to allow accessibility for cleaning on all sides, above and underneath the unit."

###  Number of Violations by borough

```{r}
#graph for the violation count by borough
pl <- df.new %>% group_by(BORO,.drop = T) %>%
        filter(BORO != "Missing" ) %>%
        summarise(total = n()) %>%
          plot_ly(x=~ BORO , 
                  y = ~total , 
                  type = "bar",
                 marker=list(color= ~total , size=20 , opacity=0.9)) %>% layout(xaxis = list(title = "Borough"),
                                                yaxis = list(title = "Number of Violations"))

pl
```

Data Analysis 
====================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Violations by Weekday and Borough

````{r}
#extract the Days of the week and agragate it to a column on the dataset
df.new <- df.new %>% mutate(DAY = format(`INSPECTION DATE`, "%A"))

#group and get the total of count of week and borough exclude some "Missing" values
total.weekday.borough <- df.new %>% group_by(BORO,DAY) %>% summarise(total = n()) %>% filter(BORO != "Missing" ) 


Week.days=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

#create a ggplot grapgh
gplot <- ggplot(total.weekday.borough, aes(x=BORO, y=total)) + geom_bar(stat = "identity", aes(fill=ordered(DAY, levels =Week.days), name = "Weekdays"), position ="dodge") + labs(x= "Borough", y="Number of Violations", title = "Violation by weekday and Borough")

#convert this ggplot object to a plotly object and get a dynamics graph
pl <- ggplotly(gplot)
pl
```


> On this graph we are comparing the weekday violations and Borough. we can see that Manhattan as the hightest number of violations.


### Violations by Year

````{r}
#this is the graph for the year counts
pl <- df.new %>% mutate( year = format(`INSPECTION DATE`, "%Y")) %>%
    group_by(year) %>%
        summarise(total = n()) %>%
          plot_ly(x=~ year , 
                  y = ~total , 
                  type = "bar",
                 marker=list(color= ~year , size=20 , opacity=0.5) ) %>% layout(xaxis = list(title = "Years"),yaxis = list(title = "Number of Violations"))
  

pl


```


> This is how many violations inspections happened by Year counts.we can see by looking at the graph that the numbers are increasing by year.



### Violations by Category

````{r}

category <- c("Critical", "Not Critical","Not Applicable")

#This is the graph for the category of critical or not critical
pl <- df.new %>% 
    group_by(`CRITICAL FLAG`) %>%
        summarise(total = n()) %>% na.omit() %>%
          plot_ly(x=~ ordered(`CRITICAL FLAG`, levels =category) , 
                  y = ~total , 
                  type = "bar",
                 marker=list(color= rainbow(3), size=20 , opacity=0.8) ) %>% layout(xaxis = list(title = "Category"),yaxis = list(title = "Number of Violations"))
  

pl


```





### Conclusion

The first impression of the data is that each year there are more violations reported for NYC restaurants, with violation code 10F[1] being the most common. Also, Manhattan has the most violations amongst the five boroughs, with about 40% of all reported violations taking place in Manhattan.

Some considerations for this analysis are that durring the data wrangling, there were many rows with missing or unavailable data, which were removed for the purpose of this project. Additionally, the Map Marker only has selected sample data as the full data set was very large, thus performance was affected. The NYC OpenData-DOHMH website source shows that the dataset was created on August 1, 2014, which most likely is the reason there is only visible data after 2015.



>[1] Violation code 10F is definted as "non-food contact surface improperly constructed. Unacceptable material used. Non-food contact surface or equipment improperly maintained and/or not properly sealed, raised, spaced or movable to allow accessibility for cleaning on all sides, above and underneath the unit."

Map Marker
================

Row
----------------------------

### Map

```{r}

  #reduced the dataset for the map in order to get better performance
  leaflet(data = df.new[1:20000,]) %>% addTiles() %>%
    addMarkers(~longitude, ~latitude, label= ~as.character(`CRITICAL FLAG`), popup    = ~as.character(paste("Grade: ",GRADE,"Violation Code:", `VIOLATION CODE`,"Date:",`INSPECTION DATE`)),clusterOptions = markerClusterOptions() , options = leafletOptions(preferCanvas = F) ) 

```

>For the Map markers I reduced the rows to 20,000 in order to get a better performance.

 Data Table
=========================

Row {.tabset .tabset-fade}
-------------------

### Data Table 

```{r}
#showing some of the data from the whole dataset. had to reduced the size of the data to obtain a better performace loading the data into datatable.
datatable(df.new[1:20000,])

```

### Dictionary for Violation Code

```{r}

datatable(description)
```





